Supplemental Information: Continuous developmental changes in word recognition support language learning across early childhood

### 2025-09-24 paper/supplemental_information.Rmd
assorted table futzing, refs

---
title: "Supplemental Information: Continuous developmental changes in word recognition support language learning across early childhood"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes:
  - \usepackage{threeparttablex}
  - \usepackage{booktabs}
  - \renewcommand{\thefigure}{S\arabic{figure}}
  - \renewcommand{\thetable}{S\arabic{table}}
  - \usepackage{float}
  - \renewcommand{\textfraction}{0.00}
  - \renewcommand{\topfraction}{1}
  - \renewcommand{\bottomfraction}{1}
  - \renewcommand{\floatpagefraction}{1}
  - \setcounter{topnumber}{5}
  - \setcounter{bottomnumber}{5}
  - \setcounter{totalnumber}{6}

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE)
library(here)
source(here("helper","common.R"))
library(papaja)
```

@@

@@

Eye-tracking data are stored in Peekbank as a time series of fixations to specific areas of interest (in particular, the target and distractor on each trial). Other fixations can be to areas not in the target or distractor as well to off-screen areas. This time series has a uniform sample rate of 25ms/sample, based on resampling of the data in Peekbank to 40 Hz during preprocessing [@zettersten2023peekbank]. Reaction times are computed by filtering trials to only those on which the child is fixating the distractor at the point of disambiguation ($t=0$) and then finding those trials on which the first non-missing fixation is to the target (hence excluding trials without a shift and trials on which a shift is to an off-screen location). The reaction time is then the total time from $t=0$ to the first timestep during which the child fixates the target. Consistent with standard practice in the literature following Fernald et al. (2008), RTs that are shorter than 367 ms are excluded as they are too short to be considered a response to the stimulus.

@@

# Separability of Variability from Other Measures

(TODO someone should edit this text)
A potential concern with multiple measures deriving from the same underlying data (as with RT/accuracy in the looking-while-listening paradigm) is that distributional assumptions about each variable could implicitly assume some connections between RT/accuracy and variability that might not be warranted.

To begin, we note that RT/accuracy and variability could be connected due to true age-related covariation - that is, some common cause leading to improvement in both, namely improvement in speed of response and improvement in consistency of response (e.g., skill learning). But they could be necessarily connected by virtue of their distributional form, and this is the concern we address here.

@@

#note that fitdistrplus attaches MASS which also has a select function, and that's annoying
detach("package:dplyr", character.only = TRUE)
library("dplyr", character.only = TRUE)

rt_fitting  <- d_sub |>
  ungroup() |>
  filter(!is.na(rt)) |>
  mutate(age_bin = cut(age,
                       breaks = c(0,12,15,18,21,24,27,30,36,60), 
                       labels = c("<12", "12-15","15-18", "18-21", 
                                  "21-24", "24-27","27-30","30-36", ">36"))) |>
  select(age_bin, rt)

@@

#BIC(norm)
#BIC(lnorm)
#BIC(wald)
#BIC(exgaus)

rt_fits <- rt_fitting |>
  group_by(age_bin) |>
  nest(rt) |>
  mutate(
    norm = map(data, \(d) fitdist(d$rt, distr = "norm", method = "mle")),
    lnorm = map(data, \(d) fitdist(d$rt, distr = "lnorm", method = "mle")),
    wald = map(data, \(d) fitdist(d$rt, distr = "invgauss", method = "mle", 
                                  start = list(mean = 1, shape = 1, dispersion = 1))),
    exgaus = map(data, \(d) fitdist(d$rt, distr = "exGAUS", method = "mle", 
                                    start = list(mu = 1000, sigma = 500, nu = 1), 
                                    lower = c(0, 0, 0)))) |>
  select(-data) |>
  pivot_longer(norm:exgaus, names_to = "distribution", values_to = "fit") |>
  mutate(BIC = map_dbl(fit, BIC)) |>
  group_by(age_bin) |>
  mutate(min_bic = BIC == min(BIC))
```

@@

We’ll look at reaction time first. Under certain distributional forms for reaction time, there would be a necessary relationship between RT and its variation. However, the literature focuses on the use of the Exponential-Gaussian (ex-Gaussian) distribution as well as Wald, Weibull, gamma, and log normal distributions (see for example Luce, 1986; Ratcliff, 1993; Van Zandt, 2002). All of these are 2- or 3-parameter distributions meaning that there is no necessary relationship between mean and variance.

@@

That said, we have attempted some fitting and we find that overall the data are best fit by either an ex-Gaussian or a log normal distribution, again consistent with prior literature, giving us confidence in this conclusion. Across the full dataset, the BIC values for ex-Gaussian (`r BIC(exgaus)`) and log normal (`r BIC(lnorm)`) are quite close to one another, and better than the Wald (`r BIC(wald)`) and normal (`r BIC(norm)`) fits. When binned by age (Fig \ref{fig:rt-fits}), younger children seem better fit for a log normal distribution and older children seem better fit by ex-Gaussian (models with lowest BICs are shown in red since significant differences can be obscured by the large scale). Figure \ref{fig:rt-hist} shows the RT data distribution overlaid with the corresponding log-normal distributions.

Overall, we think this result generally vitiates our decision to use log-transformed RTs as our primary dependent measure, and further suggests that reaction time means and variabilities are, in principle, theoretically independent.

```{r rt-fits, fig.cap="\\label{fig:rt-fits} Goodness of fit for different distributional models for RT, split by age.", fig.pos="h!", fig.width=8, fig.height=6}
ggplot(rt_fits, 
       aes(x = distribution, y = BIC, col = min_bic)) + 
  geom_point(stat = "identity", position = "dodge") + 
  scale_color_solarized(name = "Lowest BIC") + 
  xlab("Age bin") + 
  ylab("BIC") + 
  coord_flip() + 
  facet_wrap(~age_bin, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r rt-hist, fig.cap="\\label{fig:rt-hist} Distribution of RT overlaid with log normal distribution, split by age.", fig.pos="h!", fig.width=8, fig.height=6}

@@

acc_fitting  <- d_sub |>
  ungroup() |>
  filter(!is.na(long_window_accuracy)) |>
  mutate(age_bin = cut(age,
                       breaks = c(0,12,15,18,21,24,27,30,36,60), 
                       labels = c("<12", "12-15","15-18", "18-21", 
                                  "21-24", "24-27","27-30","30-36", ">36"))) |>
  select(age_bin, long_window_accuracy)

@@

#BIC(norm)
#BIC(beta)
```

Individual trial-level accuracies are not binomial because they are an average probability of fixation over a viewing window, as described in the manuscript. They are bounded at 0 and 1, but in general they tend towards the range .5 - .8 in most studies of this population. Figure \ref{acc-hist} shows the data binned by age group with fitted gaussian distributions.

These distributions seem well fit by standard gaussians, but they are in principle bounded and so we asked whether this made a difference, using a Beta distribution (a two-parameter continuous distribution bounded at 0 and 1) for fitting. Surprisingly, across all data, the BIC values for the two distributions were very similar (`r BIC(norm_acc)` for normal versus `r BIC(beta_acc)` for Beta), though the Beta was slightly favored. Across age groups, there was heterogeneity with some groups better fit by a gaussian and others better fit by a Beta (Figure \ref{fig:acc-fits}).

Again, we feel that this result generally vitiates our approach of modeling accuracies via standard linear mixed effects models: their distributional form is quite close to normal. Thus, again there is no necessary connection between mean and variance: they could in principle be independent. As above, we will plan on including this information in the Supplement.

```{r acc-hist, fig.cap="\\label{fig:acc-hist} Goodness of fit for different distributional models of accuracy, split by age.", fig.pos="h!", fig.width=8, fig.height=6}
acc_hist_data <- d_sub |>
  mutate(
    age_bin = cut(age,
                  breaks = c(0,12,15,18,21,24,27,30,36,60), 
                  labels = c("<12", "12-15","15-18", "18-21", 
                             "21-24", "24-27","27-30","30-36", ">36")))

@@

```{r acc-fits, fig.cap="\\label{fig:acc-fits} Distribution of accuracies overlaid with normal distribution, split by age.", fig.pos="h!", fig.width=8, fig.height=6}
acc_fits <- acc_fitting |>
  group_by(age_bin) |>
  nest(long_window_accuracy) |>
  mutate(
    norm = map(data, \(d) fitdist(d$long_window_accuracy, distr = "norm", method = "mle")),
    beta = map(data, \(d) fitdist(d$long_window_accuracy, distr = "beta", method = "mle"))) |>
  select(-data) |>
  pivot_longer(norm:beta, names_to = "distribution", values_to = "fit") |>
  mutate(BIC = map_dbl(fit, BIC)) |>
  group_by(age_bin) |>
  mutate(min_bic = BIC == min(BIC))

@@

# Correlations

@@

```{=tex}
\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\begin{tabular}{llllllllll}
\toprule
& \multicolumn{1}{l}{age} & \multicolumn{1}{l}{log age} & \multicolumn{1}{l}{RT} & \multicolumn{1}{l}{log RT} & \multicolumn{1}{l}{long acc} & \multicolumn{1}{l}{short acc } & \multicolumn{1}{l}{prod} & \multicolumn{1}{l}{comp}\\
\midrule
age &  1.00 &  &  &  &  &  &  & \\
log age  & 0.98 & 1.00 &  &  &  &  &  &\\
rt  & -0.33 & -0.35 & 1.00 &  &  &  &  & \\
log rt  & -0.34 & -0.36 & 0.96 & 1.00 &  &  &  & \\
long window accuracy  & 0.44 & 0.48 & -0.48 & -0.46 & 1.00 &  &  & \\
short window accuracy & 0.38 & 0.43 & -0.62 & -0.61 & 0.82 & 1.00 &  & \\
production vocabulary & 0.72 & 0.70 & -0.31 & -0.33 & 0.51 & 0.45 & 1.00 & \\
comprehension vocabulary & 0.42 & 0.42 & -0.25 & -0.24 & 0.24 & 0.24 & 0.59 & 1.00\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\caption{Pairwise correlations between primary variables of interest. \label{tab:corrs}}
\end{table}
```
```{r corrs, results = "asis", include = FALSE}

cor(select(d_sub, age, log_age, rt, log_rt, 
           long_window_accuracy, short_window_accuracy, prod, comp), 
    use = "pairwise.complete.obs") |>
  round(2) |>
  papaja::apa_table()

@@

```{=tex}
\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\begin{tabular}{lllllllll}
\toprule
n. obs & \multicolumn{1}{c}{sigma} & \multicolumn{1}{c}{logLik} & \multicolumn{1}{c}{AIC} & \multicolumn{1}{c}{BIC} & \multicolumn{1}{c}{REMLcrit} & \multicolumn{1}{c}{df.residual} & \multicolumn{1}{c}{model} & \multicolumn{1}{c}{$r_2$}\\
\midrule
48354 & 0.28 & -8,338.83 & 16,691.65 & 16,753.16 & 16,677.65 & 48347 & Long window, linear age & 0.14\\
48354 & 0.28 & -8,305.39 & 16,624.78 & 16,686.28 & 16,610.78 & 48347 & Long window, log age & 0.12\\
50244 & 0.31 & -12,486.60 & 24,987.21 & 25,048.98 & 24,973.21 & 50237 & Short window, linear age & 0.09\\
50244 & 0.31 & -12,458.77 & 24,931.54 & 24,993.31 & 24,917.54 & 50237 & Short window, log age & 0.08\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\caption{Model comparison metrics for different functional forms of the relationship between accuracy and age. \label{tab:accmods}}
\end{table}
```
```{=tex}
\begin{table}[tbp]
\begin{center}
\begin{threeparttable}
\begin{tabular}{lllllllll}
\toprule
n. obs & \multicolumn{1}{c}{sigma} & \multicolumn{1}{c}{logLik} & \multicolumn{1}{c}{AIC} & \multicolumn{1}{c}{BIC} & \multicolumn{1}{c}{REMLcrit} & \multicolumn{1}{c}{df.residual} & \multicolumn{1}{c}{model} & \multicolumn{1}{c}{$r_2$}\\
\midrule
18940 & 0.45 & -12,414.69 & 24,843.38 & 24,898.32 & 24,829.38 & 18933 & Log RT, linear age & 0.21\\
18940 & 0.45 & -12,393.27 & 24,800.53 & 24,855.48 & 24,786.53 & 18933 & Log RT, log age & 0.21\\
18940 & 570.49 & -147,644.43 & 295,302.87 & 295,357.81 & 295,288.87 & 18933 & Linear RT, linear age & 0.18\\
18940 & 570.02 & -147,622.67 & 295,259.34 & 295,314.29 & 295,245.34 & 18933 & Linear RT, log age & 0.18\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}
\caption{Model comparison metrics for different functional forms of the relationship between RT and age. \label{tab:rtmods}}
\end{table}
```
```{r accmods, results="asis", include = FALSE}
papaja::apa_table(acc_mods_lmer_summary)
```

```{r results="asis", include = FALSE}
papaja::apa_table(rt_mods_lmer_summary)
```

@@

In the literature on the "law of practice", although the log-log relationship we observed is commonly present in the aggregate across individuals, the situation is substantially more complex when relationships are measured within individuals. The best fitting curves for individuals are often exponentials or delayed exponentials (Evans et al., 2018; Heathcote, Brown, & Mewhort, 2000). With our current dataset we cannot specifically determine whether within-individual patterns of change conform to linear, power law, or exponential developmental patterns unfortunately, because we have insufficient data about individuals' improvement across time. Thus, our current results apply to the form of the age gradient as opposed to the form of any individual's pattern of developmental change.

Nevertheless, here we test for other forms of the aggregate relationship between age and reaction time. In particular, we consider further model comparison between the

@@

rt_more_mods_lmer_summary <- map_df(rt_more_mods_lmer, ~broom.mixed::glance(.x)) |>
  mutate(model = names(rt_more_mods_lmer), 
         r2 = map_dbl(rt_more_mods_lmer, ~performance::r2_nakagawa(.x)$R2_conditional))
```

@@

Here we provide specifications for the mixed effects models used in the main text. These models are used to estimate the relationship between age and the primary variables of interest, controlling for dataset and subject-level variability.

<!--\verbatim{lmer(long_window_acc_var ~ log_age_s + (log_age_s | dataset_name) + (1 | subject_id)}-->

FIXME because I cause things to crash!

@@

```{r parallel, fig.cap="Parallel analysis scree plot showing the eigenvalues for each factor, for actual, simulated, and resampled data."}
d_sub_mat <- d_sub |>
  ungroup() |>
  select(dataset_name, rt, rt_var, long_window_accuracy, long_window_acc_var, prod, comp, age)

d_sub_mat_s <- d_sub_mat |>
  ungroup() |>
  mutate(across(all_of(c("rt", "rt_var", "long_window_accuracy", 
                         "long_window_acc_var", "prod", "comp")), 
                ~ age_scale(.x, age))) 
outputs <- capture.output(fa.parallel(select(d_sub_mat, -dataset_name, -age), fa = "fa", 
                                      use = "pairwise.complete.obs"))
```

```{r eval=FALSE}
fs <- fa(select(d_sub_mat, -dataset_name), nfactor = 3,
         use = "pairwise", rotate = "varimax")

@@

```{=tex}
\begin{table}[ht]
\centering
\begin{tabular}{rrrr}
\hline
& F1 & F2 & F3 \\ 
\hline
RT & -0.19 & 0.81 & -0.30 \\ 
RT var & -0.10 & 0.81 & -0.22 \\ 
long window accuracy & 0.33 & -0.31 & 0.55 \\ 
long window accuracy var & -0.10 & 0.26 & -0.65 \\ 
production vocabulary & 0.95 & -0.04 & 0.30 \\ 
comprehension vocabulary & 0.61 & -0.16 & -0.01 \\ 
age & 0.63 & -0.14 & 0.37 \\ 
\hline
\end{tabular}
\caption{\label{tab:loadings}Factor loadings for the exploratory three factor solution using varimax rotation.} 
\end{table}
```
# Factor Analysis: First Administrations

```{r}
d_sub_mat_firstadmin <- d_sub_firstadmin |>
  ungroup() |>
  select(dataset_name, log_rt, log_rt_var, long_window_accuracy, long_window_acc_var, prod, comp, age, log_age)

>>>>

---
title: "Supplemental Information: Continuous developmental changes in word recognition support language learning across early childhood"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes:
  - \usepackage{threeparttablex}
  - \usepackage{booktabs}
  - \usepackage{placeins}
  - \renewcommand{\thefigure}{S\arabic{figure}}
  - \renewcommand{\thetable}{S\arabic{table}}
  - \usepackage{float}
  - \renewcommand{\textfraction}{0.00}
  - \renewcommand{\topfraction}{1}
  - \renewcommand{\bottomfraction}{1}
  - \renewcommand{\floatpagefraction}{1}
  - \setcounter{topnumber}{5}
  - \setcounter{bottomnumber}{5}
  - \setcounter{totalnumber}{6}

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE)
library(here)
source(here("helper","common.R"))
#library(kableExtra)
library(papaja)
```

@@

\FloatBarrier

@@

Eye-tracking data are stored in Peekbank as a time series of fixations to specific areas of interest (in particular, the target and distractor on each trial). Other fixations can be to areas not in the target or distractor as well to off-screen areas. This time series has a uniform sample rate of 25ms/sample, based on resampling of the data in Peekbank to 40 Hz during preprocessing (Zettersten et al. 2023). Reaction times are computed by filtering trials to only those on which the child is fixating the distractor at the point of disambiguation ($t=0$) and then finding those trials on which the first non-missing fixation is to the target (hence excluding trials without a shift and trials on which a shift is to an off-screen location). The reaction time is then the total time from $t=0$ to the first timestep during which the child fixates the target. Consistent with standard practice in the literature following Fernald et al. (2008), RTs that are shorter than 367 ms are excluded as they are too short to be considered a response to the stimulus.

\FloatBarrier

@@

\FloatBarrier

# Checks on data distributional assumptions

Here, we check whether the distributional forms that are assumed for the distributions of RT and accuracy are a reasonable empirical fit to the data, and compare against other commonly used distributional forms.

We confirm that, across the age range, the choice to use a log-normal distribution for RT and a normal distribution for accuracy is justified.

@@

#note that fitdistrplus attaches MASS which also has a select function, and that's annoying

rt_fitting  <- d_sub |>
  ungroup() |>
  filter(!is.na(rt)) |>
  mutate(age_bin = cut(age,
                       breaks = c(0,12,15,18,21,24,27,30,36,60), 
                       labels = c("<12", "12-15","15-18", "18-21", 
                                  "21-24", "24-27","27-30","30-36", ">36"))) |>
  dplyr::select(age_bin, rt)

@@

norm_bic <- BIC(norm)
lnorm_bic <- BIC(lnorm)
wald_bic <- BIC(wald)
exgaus_bic <- BIC(exgaus)

rt_fits <- rt_fitting |>
  group_by(age_bin) |>
  nest(rt) |>
  mutate(
    norm = map(data, \(d) fitdist(d$rt, distr = "norm", method = "mle")),
    lnorm = map(data, \(d) fitdist(d$rt, distr = "lnorm", method = "mle")),
    wald = map(data, \(d) fitdist(d$rt, distr = "invgauss", method = "mle", 
                                  start = list(mean = 1, shape = 1, dispersion = 1))),
    exgaus = map(data, \(d) fitdist(d$rt, distr = "exGAUS", method = "mle", 
                                    start = list(mu = 1000, sigma = 500, nu = 1), 
                                    lower = c(0, 0, 0)))) |>
  dplyr::select(-data) |>
  pivot_longer(norm:exgaus, names_to = "distribution", values_to = "fit") |>
  mutate(BIC = map_dbl(fit, BIC)) |>
  group_by(age_bin) |>
  mutate(min_bic = BIC == min(BIC))
```

@@

The literature focuses on the use of the Exponential-Gaussian (ex-Gaussian) distribution as well as Wald, Weibull, gamma, and log normal distributions (see for example Luce, 1986; Ratcliff, 1993; Van Zandt, 2002). All of these are 2- or 3-parameter distributions meaning that there is no necessary relationship between mean and variance.

@@

That said, we find that overall the data are best fit by either an ex-Gaussian or a log normal distribution, again consistent with prior literature, giving us confidence in this conclusion. Across the full dataset, the BIC values for ex-Gaussian (`r round(exgaus_bic)|> as.character()`) and log normal (`r round(lnorm_bic)|> as.character()`) are quite close to one another, and better than the Wald (`r round(wald_bic)|> as.character()`) and normal (`r round(norm_bic)|> as.character()`) fits. When binned by age (Fig \ref{fig:rt-fits}), younger children seem better fit for a log normal distribution and older children seem better fit by ex-Gaussian (models with lowest BICs are shown in red since significant differences can be obscured by the large scale). Figure \ref{fig:rt-hist} shows the RT data distribution overlaid with the corresponding log-normal distributions.

Overall, we think this result generally vitiates our decision to use log-transformed RTs as our primary dependent measure.

```{r rt-fits, fig.cap="\\label{fig:rt-fits} Goodness of fit for different distributional models for RT, split by age.", fig.pos="h!", fig.width=8, fig.height=5}
ggplot(rt_fits, 
       aes(x = distribution, y = BIC, col = min_bic)) + 
  geom_point(stat = "identity", position = "dodge") + 
  scale_color_solarized(name = "Lowest BIC") + 
  xlab("Age bin") + 
  ylab("BIC") + 
  coord_flip() + 
  facet_wrap(~age_bin, scales = "free_y") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r rt-hist, fig.cap="\\label{fig:rt-hist} Distribution of RT overlaid with log normal distribution, split by age.", fig.pos="h!", fig.width=8, fig.height=5}

@@

acc_fitting  <- d_sub |>
  ungroup() |>
  filter(!is.na(long_window_accuracy)) |>
  mutate(age_bin = cut(age,
                       breaks = c(0,12,15,18,21,24,27,30,36,60), 
                       labels = c("<12", "12-15","15-18", "18-21", 
                                  "21-24", "24-27","27-30","30-36", ">36"))) |>
  dplyr::select(age_bin, long_window_accuracy)

@@

norm_acc_bic <- BIC(norm_acc)
beta_acc_bic <- BIC(beta_acc)
```

Individual trial-level accuracies are not binomial because they are an average probability of fixation over a viewing window. They are bounded at 0 and 1, but in general they tend towards the range .5 - .8 in most studies of this population. Figure \ref{fig:acc-hist} shows the data binned by age group with fitted gaussian distributions.

These distributions seem well fit by standard gaussians, but they are in principle bounded and so we asked whether this made a difference, using a Beta distribution (a two-parameter continuous distribution bounded at 0 and 1) for fitting. Surprisingly, across all data, the BIC values for the two distributions were very similar (`r round(norm_acc_bic) |> as.character()` for normal versus `r round(beta_acc_bic)` for Beta), though the normal distribution was slightly favored. Across age groups, there was heterogeneity with some groups better fit by a gaussian and others better fit by a Beta (Figure \ref{fig:acc-fits}).

Again, we feel that this result generally vitiates our approach of modeling accuracies via standard linear mixed effects models: their distributional form is quite close to normal.

```{r acc-hist, fig.cap="\\label{fig:acc-hist} Goodness of fit for different distributional models of accuracy, split by age.", fig.pos="h!", fig.width=8, fig.height=5}
acc_hist_data <- d_sub |>
  mutate(
    age_bin = cut(age,
                  breaks = c(0,12,15,18,21,24,27,30,36,60), 
                  labels = c("<12", "12-15","15-18", "18-21", 
                             "21-24", "24-27","27-30","30-36", ">36")))

@@

```{r acc-fits, fig.cap="\\label{fig:acc-fits} Distribution of accuracies overlaid with normal distribution, split by age.", fig.pos="h!", fig.width=8, fig.height=5}
acc_fits <- acc_fitting |>
  group_by(age_bin) |>
  nest(long_window_accuracy) |>
  mutate(
    norm = map(data, \(d) fitdist(d$long_window_accuracy, distr = "norm", method = "mle")),
    beta = map(data, \(d) fitdist(d$long_window_accuracy, distr = "beta", method = "mle"))) |>
  dplyr::select(-data) |>
  pivot_longer(norm:beta, names_to = "distribution", values_to = "fit") |>
  mutate(BIC = map_dbl(fit, BIC)) |>
  group_by(age_bin) |>
  mutate(min_bic = BIC == min(BIC))

@@

\FloatBarrier
# Correlations

@@

```{=tex}
\begin{table}[H]
\caption{Pairwise correlations between primary variables of interest. \label{tab:corrs}}
\begin{center}
\begin{threeparttable}
\begin{tabular}{llllllllll}
\toprule
& \multicolumn{1}{l}{age} & \multicolumn{1}{l}{log age} & \multicolumn{1}{l}{RT} & \multicolumn{1}{l}{log RT} & \multicolumn{1}{l}{long acc} & \multicolumn{1}{l}{short acc } & \multicolumn{1}{l}{prod} & \multicolumn{1}{l}{comp}\\
\midrule
age &  1.00 &  &  &  &  &  &  & \\
log age  & 0.98 & 1.00 &  &  &  &  &  &\\
rt  & -0.33 & -0.35 & 1.00 &  &  &  &  & \\
log rt  & -0.34 & -0.36 & 0.96 & 1.00 &  &  &  & \\
long window accuracy  & 0.44 & 0.48 & -0.48 & -0.46 & 1.00 &  &  & \\
short window accuracy & 0.38 & 0.43 & -0.62 & -0.61 & 0.82 & 1.00 &  & \\
production vocabulary & 0.72 & 0.70 & -0.31 & -0.33 & 0.51 & 0.45 & 1.00 & \\
comprehension vocabulary & 0.42 & 0.42 & -0.25 & -0.24 & 0.24 & 0.24 & 0.59 & 1.00\\
\bottomrule
\end{tabular}
\end{threeparttable}
\end{center}

\end{table}
```

```{r corrs, results = "asis", include = FALSE, eval=F}

cor(dplyr::select(d_sub, age, log_age, rt, log_rt, 
           long_window_accuracy, short_window_accuracy, prod, comp), 
    use = "pairwise.complete.obs") |>
  round(2) |> 
  papaja::apa_table()

@@

```{r}
acc_mods_lmer_summary |>
  mutate(BIC=round(BIC),
         AIC=round(AIC),
         logLik=round(logLik),
                  REMLcrit=round(REMLcrit),
         model=str_replace(model, "lwa", "Long window") |> str_replace("swa", "Short window") |> str_replace("lin", "linear age") |> str_replace("log", "log age") |> str_replace("_", ", ")
         ) |>rename(`n. obs`=nobs) |>  dplyr::select(model, everything()) |> 
kable(digits=3, caption="Model comparison metrics for different functional forms of the relationship between accuracy and age. \\label{tab:accmods}", latex_options = "HOLD_position")
```

```{r}
rt_mods_lmer_summary |>
  mutate(BIC=round(BIC),
         AIC=round(AIC),
         logLik=round(logLik),
         REMLcrit=round(REMLcrit),
         model=str_replace(model, "log_", "Log RT, ") |> str_replace("lin_", "Linear RT, ") |> str_replace("lin", "linear age") |> str_replace("log", "log age") |> str_replace("_", ", ")
         ) |> rename(`n.  obs`=nobs) |> dplyr::select(model, everything()) |> 
kable(digits=3, caption="Model comparison metrics for different functional forms of the relationship between RT and age. \\label{tab:rtmods}", format="latex", booktabs=T, latex_options = "HOLD_position")
```

\FloatBarrier

@@

In the literature on the "law of practice", although the log-log relationship we observed is commonly present in the aggregate across individuals, the situation is substantially more complex when relationships are measured within individuals. The best fitting curves for individuals are often exponentials or delayed exponentials (Evans et al., 2018; Heathcote, Brown, & Mewhort, 2000).

With our current dataset we cannot specifically determine whether within-individual patterns of change conform to linear, power law, or exponential developmental patterns unfortunately, because we have insufficient data about individuals' improvement across time. Thus, our current results apply to the form of the age gradient as opposed to the form of any individual's pattern of developmental change.

We believe that, unlike the skills being studied in the prior adult literature (e.g., Anderson, 1982; Heathcote, Brown, & Mewhort, 2000; Logan, 1988), language processing is being learned over the course of a child’s lifetime. Thus, we do not expect to see within-paradigm changes in learning in what is a narrow period of time compared to the duration over which language processing skills are refined.

Nevertheless, here we test for other forms of the aggregate relationship between age and reaction time. In particular, we consider 1) a log~log relationship between RT and age (presented in the main text), 2) using both a log age and a linear age to predict log RT, 3) a quadratic relationship between age and RT, and 4) a cubic relationship between age and RT. As shown in Table \ref{tab:age-rt}, the model with a linear age term in addition to a log age term has the best fit, although the linear age term coefficient is only marginally significant (coefficients in Table \ref{tab:loglin}).

These models reveal a small but significant additional linear age term over and above log age, but – because individual participant-level fits are not possible – this term can’t really be used to weigh in on the debate about the precise nature of the learning pattern.

@@

 map_df(rt_more_mods_lmer, ~broom.mixed::glance(.x)) |>
  mutate(model = names(rt_more_mods_lmer), 
         model = case_when(
           model=="log_log" ~ "Log RT, log age",
           model=="log_log_plus"~ "Log RT, log+linear age",
           model=="lin_poly2" ~ "Linear RT, poly(age,2)",
           model=="lin_poly3" ~ "Linear RT, poly(age,3)"
         ),
         r2 = map_dbl(rt_more_mods_lmer, ~performance::r2_nakagawa(.x)$R2_conditional)) |> mutate(AIC=round(AIC), BIC=round(BIC), logLik=round(logLik),          REMLcrit=round(REMLcrit),
) |> dplyr::select(model, everything()) |>  rename(`n. obs`=nobs) |> kable(digits=3, caption="Goodness of fit comparison between different models of the relationship between age and RT. \\label{tab:age-rt}", padding=0, latex_options = "HOLD_position")
```

```{r}
 log_log_plus = lmer(log_rt ~ log_age_s + age_s +  
                             (log_age_s + age_s | dataset_name) + (1 | subject_id), 
                           data = d_trial)

summary(log_log_plus)$coef |> as_tibble(rownames="Term") |> 
  mutate(Term=case_when(
  Term=="(Intercept)"~"Intercept",
  Term=="log_age_s"~ "log age",
  Term=="age_s"~"linear age",)) |> rename(`$p$-value`=`Pr(>|t|)`)|> kable(digits=3, caption="Fixed effects coefficients for a model predicting log RT from both log age and linear age. \\label{tab:loglin}", format="latex", escape=F, booktabs=T, latex_options = "HOLD_position") |>   kableExtra::kable_styling(full_width = F)
```

\FloatBarrier

@@

Here we provide specifications for the lmer mixed effects models used in the main text. These models are used to estimate the relationship between age and the primary variables of interest, controlling for dataset and subject-level variability.

For accuracy, 4 models were run, crossing long and short windows as the dependent variable with age or log age as the predictor.

\begin{verbatim}
long_window_accuracy ~ log_age_s + (log_age_s | dataset_name) + (1 | subject_id)
long_window_accuracy ~ log_age_s + (log_age_s | dataset_name) + (1 | subject_id)
short_window_accuracy ~ age_s  + (age_s | dataset_name) + (1 | subject_id)
short_window_accuracy ~ log_age_s + (log_age_s | dataset_name) + (1 | subject_id)
\end{verbatim}

For reaction time, 4 models were run, crossing rt and log rt as the dependent variable  with age or log age as the predictor.

\begin{verbatim}
log_rt ~ age_s + (age_s | dataset_name) + (1 | subject_id)
log_rt ~ log_age_s + (log_age_s | dataset_name) + (1 | subject_id)
rt ~ age_s + (age_s | dataset_name) + (1 | subject_id)
rt ~ log_age_s + (log_age_s | dataset_name) + (1 | subject_id)
\end{verbatim}

To look at the relationship between variance in the accuracy and reaction time measures and children's age, we ran two models.

\begin{verbatim}
long_window_acc_var ~ log_age_s + (log_age_s | dataset_name) +  (1 | subject_id)
log_rt_var ~ log_age_s + (log_age_s | dataset_name) + (1 | subject_id)

\end{verbatim}

In the growth curve analysis, fit a mixed effects model predicting growth in vocabulary as a quadratic function of age, RT at
study initiation (t0), and their interaction, using the formula below

\begin{verbatim}
prod ~ poly(age_15,2) *rt_t0 + (age  | subject_id) + (1 | dataset_name)
\end{verbatim}

\FloatBarrier

@@

```{r parallel, fig.cap="Parallel analysis scree plot showing the eigenvalues for each factor, for actual, simulated, and resampled data."}
d_sub_mat <- d_sub |>
  ungroup() |>
  dplyr::select(dataset_name, rt, rt_var, long_window_accuracy, long_window_acc_var, prod, comp, age)

d_sub_mat_s <- d_sub_mat |>
  ungroup() |>
  mutate(across(all_of(c("rt", "rt_var", "long_window_accuracy", 
                         "long_window_acc_var", "prod", "comp")), 
                ~ age_scale(.x, age))) 
outputs <- capture.output(fa.parallel(dplyr::select(d_sub_mat, -dataset_name, -age), fa = "fa", 
                                      use = "pairwise.complete.obs"))
```

```{r eval=FALSE}
fs <- fa(dplyr::select(d_sub_mat, -dataset_name), nfactor = 3,
         use = "pairwise", rotate = "varimax")

@@

```{=tex}
\begin{table}[H]
\centering
\begin{tabular}{rrrr}
\hline
& F1 & F2 & F3 \\ 
\hline
RT & -0.19 & 0.81 & -0.30 \\ 
RT var & -0.10 & 0.81 & -0.22 \\ 
long window accuracy & 0.33 & -0.31 & 0.55 \\ 
long window accuracy var & -0.10 & 0.26 & -0.65 \\ 
production vocabulary & 0.95 & -0.04 & 0.30 \\ 
comprehension vocabulary & 0.61 & -0.16 & -0.01 \\ 
age & 0.63 & -0.14 & 0.37 \\ 
\hline
\end{tabular}
\caption{\label{tab:loadings}Factor loadings for the exploratory three factor solution using varimax rotation.} 
\end{table}
```

The confirmatory factor analysis of this three-factor solution was fit using the following specification

\begin{verbatim}
vocab =~ prod + comp
accuracy =~ long_window_accuracy + long_window_acc_var
speed =~ log_rt + log_rt_var
\end{verbatim}

The confirmatory factor analysis of the three-factor solution  with a relation to age was fit using the following specification

\begin{verbatim}
vocab =~ prod + comp
accuracy =~ acc + acc_sd
speed =~ log_rt + log_rt_sd

vocab ~ log_age
accuracy ~ log_age
speed ~ log_age
\end{verbatim}

The SEM with a linear growth curve used the following specification

\begin{verbatim}
accuracy_intercept =~ 1*acc_t0 + 1*acc_t1 + 1*acc_t2 + 1*acc_t3 + 1*acc_t4
accuracy_slope =~ 1*acc_t0 + 2*acc_t1 + 3*acc_t2 + 4*acc_t3 + 5*acc_t4
speed_intercept =~ 1*log_rt_t0 + 1*log_rt_t1 + 1*log_rt_t2 + 1*log_rt_t3 + 1*log_rt_t4
